Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-03-12 eswc-2012/phdsymp.tex
rewrite systematic paragraph
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@180 43ef1190-90fd-4dd7-a769-d0121f02b80e

Already in 1926, Ronald~Fischer noted that complex experiments can be
much more efficient than simple ones\footnote{Cited in
  http://en.wikipedia.org/wiki/Factorial\_experiment}, starting the
experimental design field. One of the simpler designs is ``fractional
factorial design'', in which several ``factors'' are studied. In terms
of SPARQL execution, the SPARQL engine is clearly a factor, but also,
for example, the nestedness of \texttt{OPTIONAL}s can be a factor, or
the number of triples in a basic graph pattern, etc. These numbers are
varied to different ``levels''. The key to understanding why this can
be efficient is that these variations need not occur in the same
experiment. Therefore, many combinations of factors can be studied by
carefully designing queries to cover different factors, and a
formalism called ``resolution'' has been developed to classify how
well this has been achieved, partly answering the question of
evaluation methodology for this part of the thesis. We should also
validate by comparing this benchmark with conclusions from existing
benchmarks.

A systematic approach based on statistical experimental design would
imply that all possible axes of variation are identified, to identify
possible contributions from different hardware platforms, operating
system constraints, backend tuning, triple store features, to SPARQL
language features and protocols. To make this benchmark feasible most
variations of parameters must be done by a software suite designed for
this purpose as the number of experiments that results will be very
large.

>>>>

Already in 1926, Ronald~Fischer noted that complex experiments can be
much more efficient than simple ones\footnote{Cited in
  http://en.wikipedia.org/wiki/Factorial\_experiment}, starting the
experimental design field. One of the simpler designs is ``fractional
factorial design'', in which several ``factors'' are studied. In terms
of SPARQL execution, the SPARQL engine is clearly a factor, but also,
for example, the nestedness of \texttt{OPTIONAL}s can be a factor, or
the number of triples in a basic graph pattern, etc. These numbers are
varied to different ``levels''. The key to understanding why this can
be efficient is that these variations need not occur in the same
experiment. Therefore, for the SPARQL language, many combinations of
factors can be studied by carefully designing queries to cover
different factors, and a formalism called ``resolution'' has been
developed to classify how well this has been achieved, partly
answering the question of evaluation methodology for this part of the
thesis. We should also validate by comparing this benchmark with
conclusions from existing benchmarks.

A systematic approach based on statistical experimental design means
developing methods to add factors (not only language features will be
factors), and then use this methodology to add and then vary all the
factors in an optimal fashion. Developing a software suite to perform
most of the experiments is a key requirement as the number of
experiments that results will be very large.
